---
title: "Assignment 3, Task 3"
author: "Catherine Takata"
date: "11/2/2020"
output: html_document
---

```{r setup, echo = TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(kableExtra)
```

#### Read in data set for local use 

```{r}
urchins <- read.csv(here("data", "urchins.csv"))
```

### Section 1
- Use a faceted histogram() and boxplot() to initially explore data through visualization. These function will help compare the distribution of sizes of purple and red urchins.  
- First, to practice "tidy" methods, convert the column names to lowercase using the clean_names() of the janitor package. 

```{r, warning=FALSE, message=FALSE}
urchins_clean <- urchins %>% 
  clean_names()
```


#### Histogram 
- Using the "tidy" dataset, we can use geom_histogram to create a histogram of the common name and size. 
- The histogram will visually represent the number of urchins in each category at each urchin diamter meausrement. 
- Using facet_wrap() will visually create a seperate group based on the two different common names. 

```{r, warning=FALSE}
ggplot(data = urchins_clean, aes(x = size, fill = common_name)) +
  geom_histogram(bins = 15, alpha = 0.5, show.legend = FALSE) + 
  facet_wrap(~common_name) +
  labs(title = "Initial Distribution of Purple and Red Urchin Based on Size",
       subtitle = "Takata, 2020",
       x = "Urchin Diameter (cm)",
       y = "Number of Urchins") +
  scale_fill_manual(values = c("mediumpurple4", "firebrick3")) +
  theme_light()
```
#### The histogram vary in size/distribution/tendency for purple and red urchins. The purple urchins seem to be postivley skewed, with a tail forming toward the opposing side. means lower with a lower tendency 



#### Boxplot 
- Using geom_boxplot() we can create a boxplot to visually represent the inital dataset. 
- Similar to the histogram, it will reveal the number of urchins per diameter. 
```{r, warning=FALSE, message=FALSE}
ggplot(data = urchins_clean, aes(x = size, fill = common_name)) +
  geom_boxplot(alpha = 0.5, show.legend = FALSE) +
  facet_wrap(~common_name) +
  labs(title = "Initial Distribution of Purple and Red Urchins Based on Size",
       subtitle = "Takata, 2020",
       x = "Urchin Diameter (cm)",
       y = "Number of Urchins") +
  scale_fill_manual(values = c("mediumpurple4", "firebrick3"))
```

### Section 1 ii.
#### Statistics Table 
- Create a summary table that contains mean, median, sd, and sample size for purple + red urchins 

```{r}
urchin_sum_stats <- urchins_clean %>% 
  group_by(common_name) %>% 
  summarize(mean = mean(size, na.rm = TRUE),
            median = median(size, na.rm = TRUE),
            sd = sd(size, na.rm = TRUE),
            sample_size = n(),
            .groups = 'drop')

kbl(urchin_sum_stats) %>% 
  kable_styling()
```

### Section 1 iii. 
The sample size is positively skewed distribution, central tendency, spread, etc.) compare between the two species of urchins (again, here we are ignoring site and treatment as a variable). Help to contextualize any differences you observe - for example, what is the actual and % difference in mean urchin diameter between the species? 


### Section 2 i. 
- Explore purple urchin counts by site and year 
- Convert data to DATE class, add columns to existing dataset with wrangled data 

#### Next, create a stored subset of the urchins data called purple_urchins
- Use filter() to isolate the observations for just the purple urchins 
- Using mutate() add a new column based on the date class 
- case_when() can be used as an "OR" statement, in this scenario, if "site" contains "an abbrviation"NAPL" OR "AQUE" OR "CARP" (etc), print it in a new mutated column. 

```{r}

urchins_clean <- urchins_clean %>% 
  mutate(date = mdy(date))

class(urchins_clean$date)

purple_urchins <- urchins_clean %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(year = year(date)) %>%
  mutate(site_full = case_when(
    site %in% "NAPL" ~ "Naples Reef",
    site %in% "AQUE" ~ "Arroyo Quemado Reef",
    site %in% "CARP" ~ "Carpinteria Reef",
    site %in% "MOHK" ~ "Mohawk Reef",
    site %in% "IVEE" ~ "Isla Vista Reef"
  )) 

```

### Section 2 ii. 
#### Summary table containing the total number of purple urchins by site + year 
- Count() will provide the number of observations per parameter

```{r}
sum_purp_urchins <- purple_urchins %>% 
  count(site_full, year)
```

#### Exploratory graph of purple urchin summary 
- Geom_col() can be used to represent the number of urchins per year 
- Like before, use facet_wrap() to distinguish between the five different sites 

```{r}
ggplot(data = sum_purp_urchins, aes(x = year, y = n, fill = site_full)) + 
  geom_col(show.legend = FALSE) +
  #geom_histogram(bins = 30, alpha = 0.5, show.legend = FALSE) + 
  facet_wrap(~site_full) + 
  labs(title = "Mean Annual Purple Urchin Population Per Site",
       subtitle = "Takata 2020",
       x = "Year",
       y = "Number of Urchins")
```

